---
title: "Investigating delay correction"
author: "Richard White"
date: "`r Sys.Date()`"
params:
  x: TRUE
  ra: TRUE
---

```{r setup, include = FALSE}
if(!is.data.frame(x)){
  stop("No params padded in")
} else {
  x <<- params$x
  ra <<- params$ra
}
```

```{r eval=TRUE, include=FALSE}
library(data.table)
library(ggplot2)

# Functions
TP <- function(var){
  sum(var=="TP",na.rm=T)
}

FP <- function(var){
  sum(var=="FP",na.rm=T)
}

TN <- function(var){
  sum(var=="TN",na.rm=T)
}

FN <- function(var){
  sum(var=="FN",na.rm=T)
}

PPV <- function(var){
  return(TP(var)/(TP(var)+FP(var)))
}

NPV <- function(var){
  return(TN(var)/(TN(var)+FN(var)))
}

SENS <- function(var){
  return(TP(var)/(TP(var)+FN(var)))
}

SPEC <- function(var){
  return(TN(var)/(TN(var)+FP(var)))
}

d <- x[,c(
  "YoDi","WoDi","nb","zscore",
  "predzscore0","predzscore1","predzscore2","predzscore3","predzscore4","predzscore5",
  "pred0","pred1","pred2","pred3","pred4","pred5",
  "WR0","WR1","WR2","WR3","WR4","WR5"
  )]
setnames(d,c("predzscore0","predzscore1","predzscore2","predzscore3","predzscore4","predzscore5"),c("pzscore0","pzscore1","pzscore2","pzscore3","pzscore4","pzscore5"))
d <- melt.data.table(d,id.vars=c("YoDi","WoDi","nb","zscore"),measure=patterns("^pzscore","^pred","^WR"))
setnames(d,c("value1","value2","value3"),c("pzscore","pnb","WR"))
d[,season:=ifelse(WoDi %in% 21:39,"Summer","Winter")]
d[,delay:=as.numeric(variable)-1]

d[,error_pnb:=pnb-nb]

d[,truePos:=zscore>2]
d[,testPos:=pzscore>2]

d[,test_results:=as.character(NA)]
d[truePos==TRUE & testPos==TRUE, test_results:="TP"]
d[truePos==FALSE & testPos==FALSE, test_results:="TN"]
d[truePos==FALSE & testPos==TRUE, test_results:="FP"]
d[truePos==TRUE & testPos==FALSE, test_results:="FN"]

```

## Traditional view

```{r echo=FALSE, fig.height=5, fig.width=7, message=FALSE, warning=FALSE}
q <- ggplot(x[wk>max(wk-50)],aes(x=wk))
q <- q + geom_ribbon(aes(ymin=-Inf,ymax=UPIb2),alpha=0.7,fill="green")
q <- q + geom_ribbon(aes(ymin=UPIb2,ymax=Inf),alpha=0.7, fill="yellow")
q <- q + geom_line(aes(y=Pnb),colour="black")
q <- q + geom_line(aes(y=nbc),colour="red")
#q <- q + scale_x_continuous("Year",breaks=seq(2000,3000,2))
#q <- q + scale_y_continuous("Proportion of deaths recorded")
#q <- q + labs(title="Proportion of deaths recorded within a certain number of days")
#q <- q + guides(fill = guide_legend(reverse = TRUE))
q
```

## Raw deaths

```{r echo=FALSE, fig.height=5, fig.width=7, message=FALSE, warning=FALSE}
setDT(ra)
ra[,delay:=as.numeric(difftime(DoR,DoD,units="days"))]
ra[,year:=as.numeric(format.Date(DoD, "%G"))]

r <- ra[,.(
  days00_06=mean(delay %in% 0:6),
  days07_13=mean(delay %in% 7:13),
  days14_20=mean(delay %in% 14:20),
  days21_27=mean(delay %in% 21:27),
  days28p=mean(delay >= 28)
),by=.(year)]
r <- melt.data.table(r,id.vars="year")

setattr(r$variable,"levels",c("0-6 days","7-13 days","14-20 days","21-27 days","28+ days"))

q <- ggplot(r,aes(x=year,y=value,fill=forcats::fct_rev(variable)))
q <- q + geom_col(alpha=0.75)
q <- q + scale_fill_brewer("",palette="Set1")
q <- q + scale_x_continuous("Year",breaks=seq(2000,3000,2))
q <- q + scale_y_continuous("Proportion of deaths recorded")
q <- q + labs(title="Proportion of deaths recorded within a certain number of days")
#q <- q + guides(fill = guide_legend(reverse = TRUE))
q
```

```{r echo=FALSE, fig.height=5, fig.width=7, message=FALSE, warning=FALSE}
toPlot <- d[,.(
  prop=mean(WR/nb)
),keyby=.(YoDi,delay)]
toPlot[,pdf:=shift(prop,type="lag"),by=YoDi]
toPlot[,pdf:=prop-pdf]
toPlot[delay==0,pdf:=prop]

q <- ggplot(toPlot,aes(x=YoDi,y=pdf,fill=forcats::fct_rev(as.factor(delay))))
q <- q + geom_col(alpha=0.75)
q <- q + scale_fill_brewer("(Iso)week delay",palette="Set1")
q <- q + scale_x_continuous("Year",breaks=seq(2000,3000,2))
q <- q + scale_y_continuous("Proportion of deaths recorded")
q <- q + labs(title="Proportion of deaths recorded in isoweek")
#q <- q + guides(fill = guide_legend(reverse = TRUE))
q
```

```{r echo=FALSE, fig.height=5, fig.width=7, message=FALSE, warning=FALSE}
q <- ggplot(d,aes(x=YoDi,y=WR/nb,group=YoDi))
q <- q + geom_boxplot()
q <- q + facet_wrap(~delay)
q <- q + scale_x_continuous("Year")
q <- q + scale_y_continuous("Proportion of deaths recorded")
q <- q + labs(title="Proportion of deaths recorded in isoweek")
q
```

\newpage 

## Raw correlations

Here we investigate the correlation between:

1. The true number of recorded deaths and the predicted deaths for lags of between 0 and 5 weeks
2. The true zscore and the predicted zscore (i.e. the z-score calculated using the predicted number of deaths) for lags of between 0 and 5 weeks

```{r echo=FALSE, fig.height=5, fig.width=7, message=FALSE, warning=FALSE}
d[,.(
  corr_nb_pnb=cor(nb,pnb,use="pairwise.complete.obs"),
  corr_nb_wr=cor(nb,WR,use="pairwise.complete.obs"),
  corr_zscore_pzscore=cor(zscore,pzscore,use="pairwise.complete.obs")
),keyby=.(delay)]
```

```{r echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
q <- ggplot(d,aes(x=pzscore,y=zscore))
q <- q + geom_point()
q <- q + stat_smooth(se=F,method="lm")
q <- q + facet_wrap(~delay)
q <- q + scale_x_continuous("Z-score using predicted number of deaths at different lags")
q <- q + scale_y_continuous("True Z-score")
q
```

```{r echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
q <- ggplot(d,aes(x=pzscore,y=zscore,colour=season))
q <- q + geom_point()
q <- q + stat_smooth(se=F,method="lm")
q <- q + facet_wrap(~delay)
q <- q + scale_x_continuous("Z-score using predicted number of deaths at different lags")
q <- q + scale_y_continuous("True Z-score")
q
```

```{r echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
q <- ggplot(d,aes(x=pnb,y=nb))
q <- q + geom_point()
q <- q + stat_smooth(se=F,method="lm")
q <- q + facet_wrap(~delay)
q <- q + scale_x_continuous("Predicted number of deaths at different lags")
q <- q + scale_y_continuous("True number of recorded deaths")
q
```

\newpage 

## Performances for excess mortality alerts

We now need to investigate the performance of the alerts. That is, if excess mortality is detected in lag 0, is this actually a true alert?

Here we list the number of true positives `tp`, true negatives `tn`, false positives `fp`, false negatives `fp`, and the positive predictive value `ppv`, negative predictive value `npv`, sensitivity `sens`, and specificity `spec`.

```{r echo=FALSE, fig.height=5, fig.width=7, message=FALSE, warning=FALSE}
results <- d[,.(
  tp=TP(test_results),
  tn=TN(test_results),
  fp=FP(test_results),
  fn=FN(test_results),
  ppv=RAWmisc::Format(PPV(test_results)*100,0),
  npv=RAWmisc::Format(NPV(test_results)*100,0),
  sens=RAWmisc::Format(SENS(test_results)*100,0),
  spec=RAWmisc::Format(SPEC(test_results)*100,0)
),keyby=.(delay)]

print(results)
```

\newpage 

## Biases

We now plot the bias (predicted number of deaths minus actual number of deaths) against the true Z-score for different lags.

```{r echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
q <- ggplot(d,aes(x=zscore,y=error_pnb))
q <- q + geom_point()
q <- q + facet_wrap(~delay)
q <- q + scale_x_continuous("True Z-Score")
q <- q + scale_y_continuous("Bias (predicted number of deaths minus actual number of deaths)")
q
```

```{r echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
q <- ggplot(d,aes(x=zscore,y=error_pnb,colour=season))
q <- q + geom_point()
q <- q + facet_wrap(~delay)
q <- q + scale_x_continuous("True Z-Score")
q <- q + scale_y_continuous("Bias (predicted number of deaths minus actual number of deaths)")
q
```

\newpage

## PZscore histograms

```{r echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
q <- ggplot(d,aes(x=pzscore))
q <- q + geom_histogram(alpha=0.7)
q <- q + facet_wrap(~delay)
q <- q + scale_x_continuous("Predicted Z-Score")
q <- q + scale_y_continuous("Histogram")
q
```
